---
title: "Assignment 3 - gg"
author:
  - 
  - 
  - 
format:
  pdf:
    toc: true
    number-sections: true
    fig-cap-location: bottom
    tbl-cap-location: top
    keep-tex: true
editor: visual
---

## Executive Summary

This report investigates whether a U.S. student applying to the Master of Computer Science program at Monash University can afford the associated costs.
The student’s parents are willing to provide full financial support, with an annual budget of USD 43,593.9.
By calculating and comparing tuition, rent, visa, and insurance expenses over two years, we estimate the total cost at approximately USD 84,700.
The results indicate that while the budget is sufficient, the choice may still impose noticeable financial pressure on the family.

## Introduction

Pursuing a graduate degree abroad is a major life decision that often comes with substantial financial considerations.
This report focuses on a student from the United States who plans to apply for the Master of Computer Science program at Monash University.
The student’s tuition and living costs will be fully supported by their parents.
The family has set a fixed annual budget of USD 43,593.9 for the two-year program.
No scholarships, personal savings, or part-time work are assumed in this analysis.
The key question is whether this budget allows the student to afford Monash without putting excessive financial pressure on the family.
We analyze multiple cost components including tuition, accommodation, visa fees, and insurance.
Data is sourced from a standardized international education cost dataset in USD.
Monash University, located in Melbourne, is known for its strong computer science program but also relatively high costs.
This report aims to assess affordability under the current assumptions and explore whether the choice is financially sustainable for the student’s family.

## Methodology

**Data Import and Validation**

This analysis based on `The Cost of International Education` dataset which was imported via `read_csv(here("data/International_Education_Costs.csv"))`.

```{r import data}
#| echo: false
#| warning: true
#| message: false

library(tidyverse)
library(here)

# Read in dataset
international_stu_cost <- read_csv(here("data/International_Education_Costs.csv"))

```

This dataset provides USD-standardized values for tuition, rent, living-cost index, visa fees, and insurance across universities, in 2025 globally. We filtered for the Master of Computer Science program at every university, verified that key cost fields were numeric, and removed any rows with missing values in those columns.

**Cost Component Calculation**

Each institution’s Total yearly Cost was calculated by summing four components:

**1. Annual Tuition** = Tuition_USD ÷ Duration_Years
**2. Annual Rent** = Rent_USD × 12 
**3. Other Living Cost** = \$15,000 × (Living_Cost_Index ÷ 100) 
**4. One-off Fees** = (Visa_Fee_USD + Insurance_USD) ÷ Duration_Years 

**Income Projection, Budget and Affordability Ratio**

We sourced the 2023 U.S. median household income of \$80,160) and projected it to 2025 using a 2.8% compound growth rate: Average Household Income in 2025 = \$80,160 \* (1 + 0.028)\^2

Assuming that each student will have half of median household income as their budget to make a comparison: Budget = Average Household Income in 2025 \* 0.5

Affordability ratio was defined by: affordability ratio = student budget/cost per year ratio \> 1 will indicate that student budget can cover the annual cost.

**Analytical Environment**

All calculation and visualization were executed in R using tidyverse package (which includes readr, dplyr, tidyr, ggplot2, tibble, purrr, et c.) and Quarto for reporting.

- Figure 1 represent each university’s affordability ratio
- Table 1 represent lists the top five most and least affordable programs.

```{r data calculation}
#| echo: false
#| warning: true
#| message: false

# 1. Calculate yearly costs
master_comsci_costs <- international_stu_cost |>
  filter( Program == "Computer Science",
          Level == "Master") |>
  mutate(
    annual_tuition = Tuition_USD / Duration_Years,
    annual_rent = Rent_USD * 12,
    other_living_cost = ( Living_Cost_Index / 100 ) * 15000,
    one_off_fees = Visa_Fee_USD + Insurance_USD, #Duration_Years,
    total_yearly_costs = annual_tuition + annual_rent + other_living_cost + one_off_fees 
      ) |>
  select(!c(Exchange_Rate, Insurance_USD, Visa_Fee_USD, Rent_USD, Living_Cost_Index, Tuition_USD, Duration_Years))

# 2. Project 2023 income to 2025 and define student budget
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
student_budget_2025 <- 0.5 * income_2025

# 3. Summarise by university and compute affordability ratio
affordability_ratio <- master_comsci_costs |>
  mutate(
    StudentBudget2025  = student_budget_2025,
    AffordabilityRatio = StudentBudget2025 / total_yearly_costs
  )
```

```{r fig-affordability}
#| echo: false
#| warning: true
#| message: false

affordability_ratio |>
  tail(30) |>
  ggplot(aes(x = reorder(University, AffordabilityRatio) , y = AffordabilityRatio)) +
    geom_col(fill = "steelblue") +
    coord_flip() +
    labs(
      title    = "Affordability Ratios for Master of Computer Science Programs",
      subtitle = "Based on 50% of projected 2025 U.S. median household income",
      x        = "University",
      y        = "Affordability Ratio"
    ) 
```

```{r table-affordability}
#| echo: false
#| warning: true
#| message: false

comparison_table <- affordability_ratio |>
  arrange(desc(AffordabilityRatio)) |>
  slice_head(n = 5) |>
  mutate(Category = "Most Affordable") |>
  bind_rows(
    affordability_ratio  |>
      arrange(AffordabilityRatio) |>
      slice_head(n = 5) |>
      mutate(Category = "Least Affordable")
  ) |>
  select(Category, University, AffordabilityRatio)

knitr::kable(
  comparison_table,
  caption = "Table 1: Top 5 Most and Least Affordable MCS Programs for U.S. Students in 2025"
)
```

