---
title: "assignment3-gg"
format: pdf
editor: visual
---

## Executive Summary

## Introduction

## Methodology

**Data Import and Validation**

This analysis based on `The Cost of International Education` dataset which was imported via `read_csv(here("data/International_Education_Costs.csv"))`.

This dataset provides USD-standardized values for tuition, rent, living-cost index, visa fees, and insurance across universities, globally. We filtered for the Master of Computer Science program at every university, verified that key cost fields were numeric, and removed any rows with missing values in those columns.

Each institution’s Total First-Year Cost was calculated by summing four components:

  **1.  Annual Tuition** = Tuition_USD ÷ Duration_Years
  **2.  Annual Rent** = Rent_USD × 12
  **3.  Other Living Cost** = $15,000 × (Living_Cost_Index ÷ 100)
  **4.  One-off Fees** = Visa_Fee_USD + Insurance_USD

We sourced the 2023 U.S. median household income ($80,160) and projected it to 2025 using a 2.8% compound growth rate:

```{r fig-affordability}
#| echo: false
#| warning: true
#| message: false

```

```{r table-affordability}
#| echo: false
#| warning: true
#| message: false

```

