---
title: "Assignment 3 - gg"
author:
  - Nutchanok Saitin
  - TZU-HSUAN Yang
  - Ruowei Li
format:
  pdf:
    toc: true
    number-sections: true
    fig-cap-location: bottom
    tbl-cap-location: top
    keep-tex: true
editor: visual
---

## Executive Summary

This report investigates whether a U.S. student applying to the Master of Computer Science program at Monash University can afford the associated costs. The student’s parents are willing to provide full financial support, with an annual budget of USD 43,593.9. By calculating and comparing tuition, rent, visa, and insurance expenses over two years, we estimate the total cost at approximately USD 84,700. The results indicate that while the budget is sufficient, the choice may still impose noticeable financial pressure on the family.

## Introduction

Pursuing a graduate degree abroad is a major life decision that often comes with substantial financial considerations. This report focuses on a student from the United States who plans to apply for the Master of Computer Science program at Monash University. The student’s tuition and living costs will be fully supported by their parents. The family has set a fixed annual budget of USD 43,593.9 for the two-year program. No scholarships, personal savings, or part-time work are assumed in this analysis. The key question is whether this budget allows the student to afford Monash without putting excessive financial pressure on the family. We analyze multiple cost components including tuition, accommodation, visa fees, and insurance. Data is sourced from a standardized international education cost dataset in USD. Monash University, located in Melbourne, is known for its strong computer science program but also relatively high costs. This report aims to assess affordability under the current assumptions and explore whether the choice is financially sustainable for the student’s family.

## Methodology

**Data Import and Validation**

This analysis based on *The Cost of International Education* dataset which was imported via `read_csv(here("data/International_Education_Costs.csv"))`.

```{r import data}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(here)

# Read in dataset
international_stu_cost <- read_csv(here( "data/International_Education_Costs.csv" ))
```

The *Cost of International Education* dataset provides five key cost components — tuition, rent, living-cost index, visa fees, and insurance — each standardized to U.S. dollars and expressed in 2025 values across global universities. We then filtered for the Master of Computer Science records, verified that each component was numeric, and confirmed there were no missing values in those fields.

```{r verify dataset}
#| echo: false
#| warning: false
#| message: false

# Filter Master of Computer Science
international_stu_cost |> filter( Program == "Computer Science",
                                  Level == "Master") |>
  head()

# Specify the cost columns
cost_cols <- c(
  "Tuition_USD", "Duration_Years", "Living_Cost_Index",
  "Rent_USD", "Visa_Fee_USD",  "Insurance_USD"
)

# Check that each column’s type is numeric
international_stu_cost |> summarise(across(all_of(cost_cols), ~ class(.)[1]))

# Check if any missing in those columns
international_stu_cost |> summarise(across(all_of(cost_cols), ~ sum(is.na(.))))
```


**Cost Component Calculation**

Each institution’s total yearly cost was calculated by summing four components:

**1. Annual Tuition** = Tuition_USD ÷ Duration_Years\
**2. Annual Rent** = Rent_USD × 12\
**3. Other Living Cost** = \$17,736[^nyu] × (Living_Cost_Index ÷ 100)\
**4. One-off Fees** = (Visa_Fee_USD + Insurance_USD) ÷ Duration_Years

[^nyu]: Derived from NYU’s *Cost of Attendance (2024–25)* “Traditional Programs” table.\
We took the 9-month total living cost (\$41,236 USD), subtracted Housing (\$23,588 USD)\
and Health Insurance (\$4,346 USD), divided by 9, then annualized to 12 months: 
$$
\frac{\$41,236 - \$23,588 - \$4,346}{9} \times 12
\;\approx\;\$17,736
$$


**Income Projection, Budget and Affordability Ratio**

We obtained the 2023 U.S. median household income (\$80,160) from the U.S. Census Bureau (U.S. Census Bureau 2024) and projected it to 2025 using a compound annual growth rate of 2.8%[^cagr]:

$$
\mathrm{HouseholdIncome}_{2025} \;=\; \$80{,}160 \times (1 + 0.028)^2
$$

[^cagr]: Compound annual growth rate calculated from FRED series **MEHOINUSA646N** (Median Household Income in the United States), rising from \$62,637 in 2013 to \$80,610 in 2023:
$$
\left(\frac{80\,610}{62\,637}\right)^{1/10} - 1 \approx 0.028 \quad(2.8\%)
$$

Assuming that each student can allocate 50 % of this projected income as their annual education budget:

$$
\mathrm{StudentBudget}_{2025} \;=\; {HouseholdIncome}_{2025} \times 0.5
$$

The affordability ratio is defined as: 

$$
\mathrm{AffordabilityRatio} \;=\; \frac{\mathrm{StudentBudget}_{2025}}{\mathrm{TotalYearlyCost}}
$$

A ratio greater than 1 indicates that the student’s annual budget covers the calculated yearly cost.

\newpage

**Analytical Environment**

All calculation and visualization were executed in R using tidyverse package (which includes readr, dplyr, tidyr, ggplot2, tibble, purrr, etc.) and Quarto for reporting.

-   @fig-affordability represent each university’s affordability ratio
-   @tbl-affordability lists the top five most and least affordable programs.

```{r data calculation}
#| echo: false
#| warning: false
#| message: false

# 1. Calculate yearly costs
master_comsci_costs <- international_stu_cost |>
  filter( Program == "Computer Science",
          Level == "Master") |>
  mutate(
    annual_tuition = Tuition_USD / Duration_Years,
    annual_rent = Rent_USD * 12,
    other_living_cost = ( Living_Cost_Index / 100 ) * 15000,
    one_off_fees = Visa_Fee_USD + Insurance_USD, #Duration_Years,
    total_yearly_costs = annual_tuition + annual_rent + other_living_cost + one_off_fees 
      ) |>
  select(!c(Exchange_Rate, Insurance_USD, Visa_Fee_USD, Rent_USD, Living_Cost_Index, Tuition_USD, Duration_Years))

# 2. Project 2023 income to 2025 and define student budget
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
student_budget_2025 <- 0.5 * income_2025

# 3. Summarise by university and compute affordability ratio
affordability_ratio <- master_comsci_costs |>
  mutate(
    StudentBudget2025  = student_budget_2025,
    AffordabilityRatio = StudentBudget2025 / total_yearly_costs
  )
```

```{r fig-affordability, fig-height=100}
#| label: fig-affordability
#| fig-cap: "Affordability Ratios for Master of Computer Science Programs"
#| echo: false
#| warning: false
#| message: false

affordability_ratio |>
  ggplot(aes(x = reorder(University, AffordabilityRatio) , y = AffordabilityRatio)) +
    geom_col(fill = "steelblue") +
    coord_flip() +
    labs(
      title    = "Affordability Ratios for Master of Computer Science Programs",
      subtitle = "Based on 50% of projected 2025 U.S. median household income",
      x        = "University",
      y        = "Affordability Ratio"
    ) 
```

```{r table-affordability}
#| label: tbl-affordability
#| tbl-cap: "Top five most and least affordable universities"
#| echo: false
#| warning: false
#| message: false

comparison_table <- affordability_ratio |>
  arrange(desc(AffordabilityRatio)) |>
  slice_head(n = 5) |>
  mutate(Category = "Most Affordable") |>
  bind_rows(
    affordability_ratio  |>
      arrange(AffordabilityRatio) |>
      slice_head(n = 5) |>
      mutate(Category = "Least Affordable")
  ) |>
  select(Category, University, AffordabilityRatio)

knitr::kable(
  comparison_table,
  caption = "Table 1: Top 5 Most and Least Affordable MCS Programs for U.S. Students in 2025"
)
```

## Result

The following figure compares the total estimated cost of pursuing a Master of Computer Science across selected universities

```{r}
#| label: fig-total-cost
#| fig-cap: "Top 30 Estimated Total Cost of Master of Computer Science"
#| warning: false
#| error: false
#| echo: false

library(tidyverse)
# Load the dataset
cost_data <- read_csv("data/International_Education_Costs.csv")

# Filter the data
cost_data <- cost_data |>
  filter(Level == "Master" & Program == "Computer Science") |>
  # columns for calcuation
  mutate(Monthly_Living_Cost = Living_Cost_Index / 100 * 1709.2,
         Tuition_PerYear = Tuition_USD / Duration_Years,
         Total_Cost = Tuition_PerYear + (Rent_USD + Monthly_Living_Cost) * 12 + Visa_Fee_USD + Insurance_USD)

# Plot with top 30 of living cost
cost_data_top <- cost_data |>
  arrange(desc(Total_Cost)) |>
  slice_head(n = 30)
# Monash Label
monash_label <- cost_data_top |>
  filter(University == "Monash University")
total_cost_plot <- ggplot(cost_data_top, 
                          aes(x = reorder(University, Total_Cost),
                              y = Total_Cost,
                              fill = University == "Monash University")) +
  geom_col()+
  geom_text(data = monash_label,
            aes(label = paste0("USD ", Total_Cost)),
            hjust = -0.1, colour = "black", size = 3.5) +
  coord_flip() +
  labs(x = "University", y = "Total Estimated Cost (USD)") +
  theme_minimal() +
  theme(legend.position = "none",
    axis.text.y = element_text(size = 7))
print(total_cost_plot)
```

@fig-total-cost shows the top 30 universities with the highest estimated total cost for completing a Master of Computer Science. Monash University ranks in the **upper-middle range**, with a total cost is **USD 57,831.6** for one academic year. Although its tuition fee is moderate, relatively high rent and living expenses in Melbourne contribute to the overall cost. These results highlight the role of rent and living expenses form the full financial burden of international education.

## Conclusion

This report evaluated the total cost of a USA student with a budget of **USD 43,593.90** studying a Master of Computer Science at Monash University. By analyzing tuition fees, rent, living expenses, visa, and insurance costs across global universities which have master of computer science.

Our analysis shows that Monash University’s tuition fee of a year is **USD 21,400** which is relatively moderate compared to U.S. and U.K. universities. The cost of living index in Melbourne (72.8) and Numbeo’s estimated monthly expenses (USD 1,244.3) support the city’s living cost as a reasonably affordable location for international students. The living cost of a year will be estimated as **14,931.6** However, monthly rent of USD 1,700 significantly increases the total cost over the an year duration, reaching an estimated **USD 20,400** for single living per year. Combined with **visa (USD 450)** and **insurance fees (USD 650)**, the total cost is **USD 57831.6**, which is **well above** the budget.

Therefore, under current assumptions, pursuing this degree at Monash would **likely impose a financial burden** unless supplementary funding options such as scholarships, shared housing, or part-time employment are secured.

We recommend that international students conduct a detailed cost breakdown when planning to study abroad, considering not only tuition fee but also living and setup costs. University like Monash remain attractive due to their academic reputation and moderate tuition fee, but total affordability depends on each student’s financial context.
