---
title: "Affordability of Master of Computer Science Programs for U.S. Students in 2025"
author:
  - Nutchanok Saitin
  - TZU-HSUAN Yang
  - Ruowei Li
format:
 pdf:
    pdf-engine: xelatex
    toc: true
bibliography: bib/references.bib
csl: bib/apa.csl
editor: visual
---

## Executive Summary

This report investigates whether a U.S. student applying to the Master of Computer Science program at Monash University can afford the associated costs. The student’s parents are willing to provide full financial support, with an annual budget of USD 43,593.9. By calculating and comparing tuition, rent and living cost expenses over 2025, we estimate the annual total cost at approximately USD 57,000. The results indicate that while the budget is sufficient, the choice may still impose noticeable financial pressure on the family.

## Introduction

Pursuing a graduate degree abroad is a major life decision that often comes with substantial financial considerations. This report focuses on a student from the United States who plans to apply for the Master of Computer Science program at Monash University. The student’s tuition and living costs will be fully supported by their parents. The family has set a annual budget of USD 43,593.9 in 2025. No scholarships, personal savings, or part-time work are assumed in this analysis. The key question is whether this budget allows the student to afford Monash without putting excessive financial pressure on the family. We analyze multiple cost components including tuition, accommodation and living cost expense. Data is sourced from a standardized international education cost dataset in USD. Monash University, located in Melbourne, is known for its strong computer science program but also relatively high costs. This report aims to assess affordability under the current assumptions and explore whether the choice is financially sustainable for the student’s family.

## Methodology

**Data Source and Preparation**

Analysis is based on the *Cost of International Education* dataset [@kaggle], imported via

`read_csv(here("data/International_Education_Costs.csv"))`.

```{r import data}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(here)

# Read in dataset
international_stu_cost <- read_csv(here( "data/International_Education_Costs.csv" ))
```

It provides program duration, tuition, rent, living cost index, visa fees and insurance (all in 2023 USD). We filtered for Master of Computer Science programs, confirmed all key cost variables were numeric with no missing values, and resolved duplicate university names by averaging affordability ratios.

```{r verify dataset}
#| echo: false
#| warning: false
#| message: false
#| output: false

# Filter Master of Computer Science
international_stu_cost |> filter( Program == "Computer Science",
                                  Level == "Master" ) |>
  head()

# Specify the cost columns
cost_cols <- c(
  "Tuition_USD", "Duration_Years", "Living_Cost_Index",
  "Rent_USD", "Visa_Fee_USD",  "Insurance_USD"
)

# Check that each column’s type is numeric
international_stu_cost |> summarise(across(all_of(cost_cols), ~ class(.)[1]))

# Check if any missing in those columns
international_stu_cost |> summarise(across(all_of(cost_cols), ~ sum(is.na(.))))

# Check duplicate name in Master of Computer Science program 
international_stu_cost |>
  filter( Program == "Computer Science",
          Level == "Master"
          ) |>
  count( University ) |>
  filter( n > 1)
```

**Cost Projection and Component Calculation**

-   **Tuition and rent:** Projected with a 4.75% forecast annual inflation rate [@inflation].

-   **Other Living Cost:** Using New York as the baseline (cost index = 100), a value of $17,736 for a master’s student (excluding housing and insurance; @cost) was applied to 2025 using each university’s living cost index.

-   **Formulas:**

    -   Annual Tuition (2025): Tuition_USD (2023) / Duration_Years × 1.0475
    -   Annual Rent (2025): Rent_USD (2023) × 12 × 1.0475
    -   Other Living Cost (2025): \$17,736 × (Living_Cost_Index / 100)

-   **Visa and insurance fees:** Excluded as they were minor (\<5%) and variable.

**Income Projection, Budget and Affordability Ratio**

The 2023 U.S. median household income of \$80,160, retrieved from U.S. Census Bureau [@2023income], was projected to 2025 at a 2.8% compound annual growth rate [@cagr]. We assumed 50% of projected income could be allocated to education:

-   Student Budget (2025): Projected Income × 0.5
-   Affordability Ratio: Student Budget / Total Yearly Cost\
    (Ratio \> 1 indicates affordability for a median U.S. household.)

```{r data calculation}
#| echo: false
#| warning: false
#| message: false

# 1. Calculate yearly costs in 2023 and check proportion of one-off fees
master_comsci_2023 <- international_stu_cost |>
  filter( Program == "Computer Science",
          Level == "Master"
          ) |>
  mutate(
    annual_tuition = Tuition_USD / Duration_Years,
    annual_rent = Rent_USD * 12,
    other_living_cost = ( Living_Cost_Index / 100 ) * 17736,
    one_off_fees = Visa_Fee_USD + Insurance_USD,
    Total_Yearly_Costs = annual_tuition + annual_rent + other_living_cost + one_off_fees,
    One_off_fees_portion = one_off_fees / Total_Yearly_Costs
      ) |>
  select(!c( Exchange_Rate, Insurance_USD, Visa_Fee_USD, Rent_USD, Living_Cost_Index, Tuition_USD, Duration_Years ))

# 2. Project 2023 costs to 2025 and define 2025 cost for each institution
master_comsci_2025 <- master_comsci_2023 |>
  mutate(
    Total_2023_Costs = annual_tuition + annual_rent,
    Total_2025_Costs = (Total_2023_Costs * 1.0475) + other_living_cost
    ) |>
  select(!c( annual_tuition,  annual_rent, other_living_cost, one_off_fees, Total_Yearly_Costs, One_off_fees_portion )
         )
    
# 3. Project 2023 income to 2025 and define student budget
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
Student_Budget_2025 <- 0.5 * income_2025

# 4. Summarise by university and compute affordability ratio
affordability_ratio <- master_comsci_2025 |>
  mutate(
    StudentBudget2025  = Student_Budget_2025,
    AffordabilityRatio = StudentBudget2025 / Total_2025_Costs
    ) |>
  group_by( University ) |>
  summarise( AffordabilityRatio = mean( AffordabilityRatio, na.rm = TRUE), # handle duplicate in university names
             .groups = "drop") |>
  arrange( AffordabilityRatio )
```

**Analytical Environment**

All analyses were performed in R using tidyverse (e.g. readr, ggplot2, knitr) [@tidyverse] and here [@here] within a Quarto workflow. Git was used for version control, and the report was rendered via LaTeX.

-   @fig-affordability shows each university’s affordability ratio.
-   @tbl-affordability summarizes the 30 most and least affordable Master of Computer Science programs.

```{r fig-affordability, fig.height=17, fig.width=13}
#| label: fig-affordability
#| fig-cap: "Affordability Ratios for Master of Computer Science Programs by University in 2025"
#| echo: false
#| warning: false
#| message: false

afford_plot <- affordability_ratio |>
  mutate(
    University = factor( University, levels = unique( University )),
    Highlight  = if_else( University == "Monash University", "Monash", "Other"))

ggplot(afford_plot, aes(x = University, 
                        y = AffordabilityRatio, 
                        fill = Highlight)) +
  geom_col() +
  scale_fill_manual(values = c("Monash" = "firebrick", 
                               "Other" = "grey80")) +
  coord_flip() +
  geom_text(
    data = afford_plot |>
      filter(University == "Monash University"),
    aes(label = "Monash University", y = AffordabilityRatio),
    hjust = -0.1,            
    color = "firebrick",
    fontface = "bold",
    size = 8
  ) +
  geom_hline(yintercept = 1, 
             linetype = "dashed", 
             color = "darkblue", 
             linewidth = 1) +
  labs(
    title = "Affordability Ratios for Master of Computer Science Programs in 2025",
    subtitle = "Based on 50% of projected 2025 U.S. median household income",
    x = "University",
    y = "Affordability Ratio" 
    ) +
  theme_minimal() +
  theme(
    axis.text.y = element_text(size = 9),
    legend.position = "none"
    )
```

```{r table-affordability}
#| label: tbl-affordability
#| tbl-cap: "Top 5 Most and Least Affordable Master of Computer Science Programs (2025 Budget)"
#| echo: false
#| warning: false
#| message: false

summary_table <- afford_plot |>
  arrange(desc( AffordabilityRatio )) |>
  mutate(Rank = row_number()) |>
  
  # Get top 5 and bottom 5
  filter(Rank <= 30 | Rank > n() - 30) |>
  mutate(Category = if_else( Rank <= 30, 
                             "Most Affordable", 
                             "Least Affordable")) |>
  select(Category, University, AffordabilityRatio, Rank)

# Print table
knitr::kable( summary_table )
```

## Result

The following figure compares the estimated affordability ratio of applying a Master of Computer Science across selected universities.

```{r}
#| label: fig-ratio
#| fig-cap: "Bottom 30 Estimated Affordability Ratio of Master of Computer Science"
#| warning: false
#| error: false
#| echo: false

# Number defined in Methodology
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
student_budget_2025 <- 0.5 * income_2025

# Filter the data
cost_data <- international_stu_cost |>
  filter(Level == "Master" & Program == "Computer Science") |>
  # columns for calcuation
  mutate(AnnualLiving_Cost = ( Living_Cost_Index / 100 ) * 17736,
         Tuition_PerYear = (Tuition_USD * (1 + 0.0475)) / Duration_Years,
         Total_Cost = AnnualLiving_Cost + Tuition_PerYear + Rent_USD * (1 + 0.0475) * 12,
         AffordabilityRatio = student_budget_2025 / Total_Cost )

# Plot with top 30 of living cost
ratio_data_top <- cost_data |>
  arrange(AffordabilityRatio) |>
  slice_head(n = 30)
# Monash Label
monash_label <- ratio_data_top |>
  filter(University == "Monash University")
ratio_plot <- ggplot(ratio_data_top, 
                          aes(x = reorder(University, AffordabilityRatio),
                              y = AffordabilityRatio,
                              fill = University == "Monash University")) +
  geom_col()+
  coord_flip() +
  labs(x = "University", y = "Affordability Ratio") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "none", # hide the fill logic
    axis.text.y = element_text(size = 7)) #adjust the font
print(ratio_plot)
```

@fig-ratio shows the bottom 30 universities with the lowest affordability ratio for applying a master degree of Computer Science in 2025. Monash University ranks in the **bottom range**, with a ratio is **0.751** for one academic year. Although its tuition fee is moderate, relatively high rent and living expenses in Melbourne contribute to the overall cost. These results highlight the role of rent and living expenses form the full financial burden of international education.

## Conclusion

This report evaluated the total cost of a USA student with a budget of **USD 42,355.90** plan to apply a master degree of Computer Science at Monash University in 2025. By analyzing tuition fees, rent, living expenses, visa, and insurance costs for year 2025 (adjust by inflation rate) across global universities which have master program of computer science.

Our analysis shows that Monash University’s tuition fee of a year is **USD 22,416.5** which is relatively moderate when compared to U.S. and U.K. universities. Based on the cost of living index in Melbourne (72.8) and the [Cost of Attendance](https://bulletins.nyu.edu/graduate/public-service/cost-attendance/cost-attendance.pdf) which released by New York University (NYU), the estimated monthly living cost is **USD 1,076** and the annual living cost will be estimated as **USD 12,912**. This is also relatively moderate compared to U.S. and U.K. universities, though it still acts as an major part of total expense. However, monthly rent of USD 1,780.75 significantly increases the total cost and decreases the affordability ratio which reaching an estimated **USD 21,369** for single living per year. The total cost is **USD 56,697.5** which has a gap **USD 14,341.6** between budget, and the affordability raio is **0.751**, which is **well above**.

Therefore, under current assumptions, applying this degree at Monash would **likely impose a financial burden** unless supplementary funding options such as scholarships, shared housing, or part-time employment are secured.

We recommend that international students conduct a detailed cost breakdown when planning to study abroad, considering not only tuition fee but also living and setup costs. University like Monash remain attractive due to their academic reputation and moderate tuition fee, but total affordability depends on each student’s financial context.

## References