---
title: "Assignment 3 - gg"
author:
  - Nutchanok Saitin
  - TZU-HSUAN Yang
  - Ruowei Li 
format:
  html:
    toc: false
    fig-cap-location: bottom
    tbl-cap-location: top
  pdf:
    toc: false
    fig-cap-location: bottom
    tbl-cap-location: top
editor: visual
---

## Executive Summary

This report evaluates whether a U.S. student, fully supported by their parents, can afford to study the Master of Computer Science program at Monash University. With an annual budget of **USD 42,355.90**, we analyze tuition, rent, living expenses, visa, and insurance costs for the year 2025. The total estimated cost is **USD 55,262**, significantly exceeding the budget. Therefore, studying at Monash would impose financial pressure unless additional support such as scholarships or shared housing is available.

## Introduction

Studying abroad is a significant life decision that also requires thorough financial planning. This report explores whether a U.S. student with a fixed annual budget of **USD 42,355.90** can afford to study the Master of Computer Science at Monash University. We assume no scholarships, savings, or part-time income. Using global datasets on tuition and living costs, we isolate Monash-specific data. Tuition is estimated at **USD 21,400**, while living costs are based on the Cost of Attendance data from NYU and Numbeo’s index for Melbourne. Rent is estimated at **USD 1,700** per month, significantly affecting affordability. Combined with visa and insurance, the total estimated cost reaches **USD 55,262**. The budget shortfall suggests financial infeasibility under current conditions. The aim is to offer financial insights for prospective students planning to study at Monash.Methodology

**Data Import and Validation**

This analysis based on `The Cost of International Education` dataset which was imported via `read_csv(here("data/International_Education_Costs.csv"))`.

```{r import data}
#| echo: false
#| warning: true
#| message: false

library(tidyverse)
library(here)

# Read in dataset
international_stu_cost <- read_csv(here("data/International_Education_Costs.csv"))

```

This dataset provides USD-standardized values for tuition, rent, living-cost index, visa fees, and insurance across universities, in 2025 globally. We filtered for the Master of Computer Science program at every university, verified that key cost fields were numeric, and removed any rows with missing values in those columns.

**Cost Component Calculation**

Each institution’s Total yearly Cost was calculated by summing four components:

**1. Annual Tuition** = Tuition_USD ÷ Duration_Years **2. Annual Rent** = Rent_USD × 12 **3. Other Living Cost** = \$15,000 × (Living_Cost_Index ÷ 100) **4. One-off Fees** = (Visa_Fee_USD + Insurance_USD) ÷ Duration_Years

**Income Projection, Budget and Affordability Ratio**

We sourced the 2023 U.S. median household income of \$80,160) and projected it to 2025 using a 2.8% compound growth rate: Average Household Income in 2025 = \$80,160 \* (1 + 0.028)\^2

Assuming that each student will have half of median household income as their budget to make a comparison: Budget = Average Household Income in 2025 \* 0.5

Affordability ratio was defined by: affordability ratio = student budget/cost per year ratio \> 1 will indicate that student budget can cover the annual cost.

**Analytical Environment**

All calculation and visualization were executed in R using tidyverse package (which includes readr, dplyr, tidyr, ggplot2, tibble, purrr, et c.) and Quarto for reporting.

-   Figure 1 represent each university’s affordability ratio
-   Table 1 represent lists the top five most and least affordable programs.

```{r data calculation}
#| echo: false
#| warning: true
#| message: false

# 1. Calculate yearly costs
master_comsci_costs <- international_stu_cost |>
  filter( Program == "Computer Science",
          Level == "Master") |>
  mutate(
    annual_tuition = Tuition_USD / Duration_Years,
    annual_rent = Rent_USD * 12,
    other_living_cost = ( Living_Cost_Index / 100 ) * 15000,
    one_off_fees = Visa_Fee_USD + Insurance_USD, #Duration_Years,
    total_yearly_costs = annual_tuition + annual_rent + other_living_cost + one_off_fees 
      ) |>
  select(!c(Exchange_Rate, Insurance_USD, Visa_Fee_USD, Rent_USD, Living_Cost_Index, Tuition_USD, Duration_Years))

# 2. Project 2023 income to 2025 and define student budget
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
student_budget_2025 <- 0.5 * income_2025

# 3. Summarise by university and compute affordability ratio
affordability_ratio <- master_comsci_costs |>
  mutate(
    StudentBudget2025  = student_budget_2025,
    AffordabilityRatio = StudentBudget2025 / total_yearly_costs
  )
```

```{r fig-affordability}
#| echo: false
#| warning: true
#| message: false

affordability_ratio |>
  tail(30) |>
  ggplot(aes(x = reorder(University, AffordabilityRatio) , y = AffordabilityRatio)) +
    geom_col(fill = "steelblue") +
    coord_flip() +
    labs(
      title    = "Affordability Ratios for Master of Computer Science Programs",
      subtitle = "Based on 50% of projected 2025 U.S. median household income",
      x        = "University",
      y        = "Affordability Ratio"
    ) 
```

```{r table-affordability}
#| echo: false
#| warning: true
#| message: false

comparison_table <- affordability_ratio |>
  arrange(desc(AffordabilityRatio)) |>
  slice_head(n = 5) |>
  mutate(Category = "Most Affordable") |>
  bind_rows(
    affordability_ratio  |>
      arrange(AffordabilityRatio) |>
      slice_head(n = 5) |>
      mutate(Category = "Least Affordable")
  ) |>
  select(Category, University, AffordabilityRatio)

knitr::kable(
  comparison_table,
  caption = "Table 1: Top 5 Most and Least Affordable MCS Programs for U.S. Students in 2025"
)
```

## Result

The following figure compares the estimated affordability ratio of applying a Master of Computer Science across selected universities

```{r}
#| label: fig-ratio
#| fig-cap: "Bottom 30 Estimated Affordability Ratio of Master of Computer Science"
#| warning: false
#| error: false
#| echo: false

# Number defined in Methodology
base_income_2023    <- 80610   # from U.S. Census Bureau
income_2025         <- base_income_2023 * (1 + 0.028)^2
student_budget_2025 <- 0.5 * income_2025

# Filter the data
cost_data <- international_stu_cost |>
  filter(Level == "Master" & Program == "Computer Science") |>
  # columns for calcuation
  mutate(AnnualLiving_Cost = ( Living_Cost_Index / 100 ) * 17736,
         Tuition_PerYear = Tuition_USD / Duration_Years,
         OneTime_fee = (Visa_Fee_USD + Insurance_USD) / Duration_Years,
         Total_Cost = AnnualLiving_Cost + Tuition_PerYear + Rent_USD * 12 + OneTime_fee,
         AffordabilityRatio = student_budget_2025 / Total_Cost )

# Plot with top 30 of living cost
ratio_data_top <- cost_data |>
  arrange(AffordabilityRatio) |>
  slice_head(n = 30)
# Monash Label
monash_label <- ratio_data_top |>
  filter(University == "Monash University")
ratio_plot <- ggplot(ratio_data_top, 
                          aes(x = reorder(University, AffordabilityRatio),
                              y = AffordabilityRatio,
                              fill = University == "Monash University")) +
  geom_col()+
  geom_text(data = monash_label,
            aes(label = paste0(AffordabilityRatio)),
            # show the ratio number
            hjust = 1.1, colour = "black", size = 3) +
  coord_flip() +
  labs(x = "University", y = "Affordability Ratio") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "none", # hide the fill logic
    axis.text.y = element_text(size = 7)) #adjust the font
print(ratio_plot)
```

@fig-ratio shows the bottom 30 universities with the lowest affordability ratio for applying a master degree of Computer Science in 2025. Monash University ranks in the **bottom range**, with a ratio is **0.771** for one academic year. Although its tuition fee is moderate, relatively high rent and living expenses in Melbourne contribute to the overall cost. These results highlight the role of rent and living expenses form the full financial burden of international education.

## Conclusion

This report evaluated the total cost of a USA student with a budget of **USD 42,355.90** studying a Master of Computer Science at Monash University. By analyzing tuition fees, rent, living expenses, visa, and insurance costs for year 2025 across global universities which have master of computer science.

Our analysis shows that Monash University’s tuition fee of a year is **USD 21,400** which is relatively moderate when compared to U.S. and U.K. universities. Based on the cost of living index in Melbourne (72.8) and the [Cost of Attendance](https://bulletins.nyu.edu/graduate/public-service/cost-attendance/cost-attendance.pdf) which released by New York University (NYU), the estimated monthly living cost is **USD 1,076** and the annual living cost will be estimated as **USD 12,912**. This is relatively moderate compared to U.S. and U.K. universities. However, monthly rent of USD 1,700 significantly increases the total cost and decreases the affordability ratio which reaching an estimated **USD 20,400** for single living per year. Combined with **visa (USD 225/year)** and **insurance fees (USD 325/year)**, the total cost is **USD 55,262** and the affordability raio is **0.771**, which is **well above** the budget.

Therefore, under current assumptions, applying this degree at Monash would **likely impose a financial burden** unless supplementary funding options such as scholarships, shared housing, or part-time employment are secured.

We recommend that international students conduct a detailed cost breakdown when planning to study abroad, considering not only tuition fee but also living and setup costs. University like Monash remain attractive due to their academic reputation and moderate tuition fee, but total affordability depends on each student’s financial context.
